# Exploratory Data Analysis

```{r include = FALSE}
library(tidyverse)
library(car)
library(stringr)
library(corrplot)
library(RColorBrewer)
```

### Load Dataset

```{r}
cleandata <- read.csv("../Data/cleandata.csv", header = TRUE) %>% 
  mutate(ovaryHolesCount = ifelse(ovaryHolesSize == "0.0",
                                  0,
                                  str_count(ovaryHolesSize, ",") + 1))


```

```{r}
table(cleandata$year, cleandata$abscissed)
```

```{r}
cleandata <- cleandata %>% drop_na()
table(cleandata_nona$year, cleandata_nona$abscissed)
```



### Data Split

```{r}
#split data into test and train sets
set.seed(1829)
split <- sample(x = nrow(cleandata),size = nrow(cleandata)*.75)
train <- cleandata[split, ]
test <- cleandata[-split, ]

#report number of obs in each
paste0("The train data has ", nrow(train), " observations.")
paste0("The test data has ", nrow(test), " observations.")
```

```{r}
write.csv(train, "../Data/train.csv", row.names = FALSE)
write.csv(test, "../Data/test.csv", row.names = FALSE)
```



### (Use Training Set ONLY)  

## Tables & Plots for Variety  

#### Diameter  

```{r}
train %>% filter(diameter < 50 & diameter > 0) %>%
  ggplot(aes(diameter, fill = factor(abscissed))) +
  geom_boxplot() +
  facet_wrap(~var) + 
  labs(title = "Diameter of Cotton Boll From Each Variety",
       x = "Diameter of Boll (mm)")
```


#### Sepal  

```{r}
train %>% filter(diameter < 50 & sepal > 0) %>%
  ggplot(aes(sepal, var, fill = var)) +
  coord_cartesian(xlim = c(0, 30)) +
  geom_boxplot() + 
  labs(title = "Diameter of Sepal Damage for Each Variety",
       subtitle = "(When Present)",
       x = "Sepal Damage Diameter (mm)",
       y = "Variety")
```

#### Bract  

```{r}
train %>% filter(diameter < 50 & bract > 0) %>%
  ggplot(aes(bract, var, fill = var)) +
  coord_cartesian(xlim = c(0, 30)) +
  geom_boxplot() + 
  labs(title = "Diameter of Bract Damage for Each Variety",
       subtitle = "(When Present)",
       x = "Diameter of Bract Damage (mm)",
       y = "Variety")
```


#### Boll Wall  

```{r}
train %>% filter(diameter < 50 & bollWall > 0) %>%
  ggplot(aes(bollWall, var, fill = var)) +
  coord_cartesian(xlim = c(0, 50)) +
  geom_boxplot() + 
  labs(title = "Diameter of Boll Wall Damage for Each Variety",
       subtitle = "(When Present)",
       x = "Diameter of Boll Wall Damage (mm)",
       y = "Variety")
```


#### Total Ovary Hole Size  

```{r}
train %>% filter(diameter < 50 & ovaryHolesSum > 0) %>%
  ggplot(aes(ovaryHolesSum, var, fill = var)) +
  coord_cartesian(xlim = c(0, 30)) +
  geom_boxplot() + 
  labs(title = "Total Diameter of Ovary Hole Size for Each Variety",
       subtitle = "(When Present)",
       x = "Total Diameter of Ovary Hole Size (mm)",
       y = "Variety")
```



#### Abcission Rate by Variety and Year  

```{r}
table <- train %>% group_by(var, year) %>% summarise(Abscised = sum(abscissed)/n(), Not_Abscised = (n() - sum(abscissed))/n())
table %>% ggplot(aes(x = var, y = Abscised, fill = factor(year))) + 
  geom_col(position = "dodge") + 
  scale_y_continuous(n.breaks = 12) + 
  labs(title = "Proportion of Abscission for Each Variety Across Years",
       x = "Variety",
       y = "Proportion of Cotton Bolls Abscised",
       fill = "Year")
```



```{r}
train %>% filter(diameter < 50 & diameter > 0) %>% ggplot(aes(x = diameter, fill = factor(abscissed))) + 
  geom_histogram() + 
  labs(title = "Distribution of Diameter for Abscised vs. Non-Abscised Cotton Bolls",
       x = "Diameter",
       y = "Count",
       fill = "Abscised") + 
  scale_fill_discrete(labels = c("No", "Yes"))


train %>% filter(diameter < 50 & diameter > 0) %>% ggplot(aes(x = diameter, fill = factor(year))) + 
  geom_histogram() + 
  labs(title = "Distribution of Diameter for Each Year",
       x = "Diameter",
       y = "Count",
       fill = "Year")
```

Why is the distribution bimodal for diameter in 2015/non-abscissed tissue?  



#### Summary Statistics of Each Variable  

```{r}
a <- summary(train$diameter)
b <- summary(train$sepal)
c <- summary(train$bract)
d <- summary(train$bollWall)
e <- summary(train$ovaryHolesCount)
f <- summary(train$ovaryHolesSum)

list(diameter = a, 
     sepal = b,
     bract = c,
     bollWall = d,
     ovaryHolesCount = e,
     ovaryHolesSum = f)
```


#### Correlation Between Variables (when > 0)   


```{r}
# look at observations > 0
filtrain <- train %>% filter(diameter < 50 & diameter > 0 &
                               sepal > 0 & 
                               bract > 0 & 
                               bollWall > 0 & 
                               ovaryHolesCount > 0 & 
                               ovaryHolesSum > 0)

# correlation matrix
correlation <- cor(filtrain[c(6:10,25)])
correlation

# colorful correlation plot
corrplot(correlation, type="upper", order="hclust",
         col=brewer.pal(n=8, name="RdYlBu"))
```

Correlations to note:  
    * Correlation between SEPAL and BRACT is 0.55168  
    
Correlation should not be a problem when modeling.  



































