---
title: "Model Plotting"
author: "John Hinic"
date: '2022-07-18'
output:
  html_document:
    toc: true
    toc_depth: 3
    theme: readable
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
library(tidyverse)
library(caret)
library(doParallel)
library(yardstick)
library(rpart)
library(rpart.plot)
library(gbm)
library(randomForest)
library(ggeffects)
library(gridExtra)
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

The following logistic model will be used for plotting:

```{r}
trainPlot <- read_csv('../Data/train.csv') %>%
  filter(diameter > 0 & diameter < 3000) %>%
  mutate(
    var = factor(var, levels = c('RF', 'B2RF', 'WRF', 'GLT')), 
    year = as.factor(year), 
    rep = as.factor(rep), 
    abscissed = as.factor(abscissed)
  ) %>%
  drop_na()

logPlot <- glm(
  abscissed ~ var + diameter + year + sepal + bract + bollWall + ovaryHolesCount + ovaryHolesSum + 
    var:year + var:diameter + var:sepal + var:bract + var:bollWall + var:ovaryHolesSum + 
    diameter:bollWall + diameter:ovaryHolesCount + 
    year:sepal + year:bract + year:ovaryHolesCount + year:ovaryHolesSum + 
    I(diameter^2) + I(bract^2) + I(bollWall^2) + I(ovaryHolesCount^2) + I(ovaryHolesSum^2),
  data = trainPlot,
  family = "binomial"
)

summary(logPlot)
```


## Basic Summary Plots

These plots are basically just summaries, looking at the sample proportions for each variety and the damage types by each variety - they are NOT based on the final logistic model

```{r}
# abscission by variety
varPlot <- glm(
  abscissed ~ var,
  data = trainPlot,
  family = "binomial"
)

varCI <- varPlot %>% 
  ggeffect(terms = "var") %>% plot() +
  labs(
    x = "Variety", 
    y = "Proportion Abscised", 
    title = "",
  ) + theme_bw()
ggsave("../plots/var_basic_summary.png", plot = varCI, width = 8, height = 6)


# damage levels by variety
p1 <- lm(bract ~ var, data = trainPlot) %>% 
  ggeffect(terms = "var") %>% plot() +
  labs(y = "Bract", x = "Variety", title='Bract Damage') +
  theme_bw()

p2 <- lm(sepal ~ var, data = trainPlot) %>% 
  ggeffect(terms = "var") %>% plot() +
  labs(y = "Sepal", x = "Variety", title='Sepal Damage') +
  theme_bw()

p3 <- lm(bollWall ~ var, data = trainPlot) %>% 
  ggeffect(terms = "var") %>% plot() +
  labs(y = "Boll Wall", x = "Variety", title='Boll Wall Damage') +
  theme_bw()

p4 <- lm(ovaryHolesSum ~ var, data = trainPlot) %>% 
  ggeffect(terms = "var") %>% plot() +
  labs(y = "Ovary Holes Sum", x = "Variety", title='Sum of Ovary Holes Damage') +
  theme_bw()

g <- arrangeGrob(p1, p2, p3, p4)
ggsave("../plots/var_dmg_summary.png", g, width = 12, height = 9)
```

## Interaction Plots

### Variety Interactions

```{r}
# variety by year
var_year <- ggeffect(logPlot, c("year", "var")) %>% plot() +
  labs(x = "Year", y = "Predicted Probability of Abscission", 
       colour = "Variety", title = '') +
  theme_bw()
ggsave("../plots/var_year.png", var_year, width = 8, height = 6)

# variety by diameter
var_diam <- ggeffect(logPlot, c("diameter [all]", "var")) %>% plot(add.data = TRUE) +
  labs(x = "Diameter", y = "Predicted Probability of Abscission", 
       colour = "Variety", title = '') +
  theme_bw()
ggsave('../plots/var_diam.png', var_diam, width = 8, height = 6)

# variety by sepal
var_sepal <- ggeffect(logPlot, c("sepal [all]", "var")) %>% plot(add.data = TRUE) +
  labs(x = "Sepal", y = "Predicted Prob. of Abscission", 
       colour = "Variety", title = 'Sepal') +
  theme_bw()

# variety by bract
var_bract <- ggeffect(logPlot, c("bract [all]", "var")) %>% plot(add.data = TRUE) +
  labs(x = "Bract", y = "Predicted Prob. of Abscission", 
       colour = "Variety", title = 'Bract') +
  theme_bw()

# variety by bollWall
var_boll <- ggeffect(logPlot, c("bollWall [all]", "var")) %>% plot(add.data = TRUE) +
  labs(x = "Boll Wall (truncated)", y = "Predicted Prob. of Abscission", 
       colour = "Variety", title = 'Boll Wall') +
  scale_x_continuous(limits = c(0, 100)) +
  theme_bw()

# variety by ovaryHolesSum
var_ovary <- ggeffect(logPlot, c("ovaryHolesSum [all]", "var")) %>% plot(add.data = TRUE) +
  labs(x = "Sum of Ovary Holes", y = "Predicted Prob. of Abscission", 
       colour = "Variety", title = 'Sum of Ovary Holes') +
  theme_bw()

var_ints <- arrangeGrob(var_sepal, var_bract, var_boll, var_ovary)
ggsave("../plots/var_grid.png", var_ints, width = 8, height = 6)
```

### Diameter Interactions

```{r}
# diameter by bollWall
diam_boll <- ggeffect(logPlot, c("diameter [all]", "bollWall [0,5,10,50,100]")) %>% plot(add.data = TRUE) +
  labs(x = "Diameter", y = "Predicted Prob. of Abscission", 
       colour = "Boll Wall Damage", title = 'Boll Wall') +
  theme_bw()

# diameter by ovaryHolesCount
diam_ovary <- ggeffect(logPlot, c("diameter [all]", "ovaryHolesCount[0,2,4]")) %>% plot(add.data = TRUE) +
  labs(x = "Diameter", y = "Predicted Prob. of Abscission", 
       colour = "Num. Ovary\nHoles", title = 'Num. Ovary Holes') +
  theme_bw()

diam_ints <- arrangeGrob(diam_boll, diam_ovary)
ggsave("../plots/diam_grid.png", diam_ints, width = 8, height = 6)


# alt mapping for diameter interactions
# diameter by bollWall
diam_boll_alt <- ggeffect(logPlot, c("bollWall [all]", "diameter [5,15,25,35]")) %>% plot(add.data = TRUE) +
  labs(x = "Boll Wall Damage", y = "Predicted Prob. of Abscission", 
       colour = "Diameter", title = 'Boll Wall') +
  theme_bw()

diam_ints_alt <- arrangeGrob(diam_boll_alt, diam_ovary)
ggsave("../plots/diam_grid_alt.png", diam_ints_alt, width = 8, height = 6)
```

### Year Interactions

```{r}
# year by sepal
year_sepal <- ggeffect(logPlot, c("sepal [all]", "year")) %>% plot(add.data = TRUE) +
  labs(x = "Sepal", y = "Predicted Prob. of Abscission",
       colour = "Year", title = 'Sepal') +
  theme_bw()

# year by bollWall
# NOT SIGNIFICANT
year_boll <- ggeffect(logPlot, c("bollWall [all]", "year")) %>% plot(add.data = TRUE) +
  labs(x = "Boll Wall", y = "Predicted Prob. of Abscission",
       colour = "Year", title = 'Boll Wall') +
  theme_bw()

# year by bract
year_bract <- ggeffect(logPlot, c("bract [all]", "year")) %>% plot(add.data = TRUE) +
  labs(x = "Bract", y = "Predicted Prob. of Abscission",
       colour = "Year", title = 'Bract') +
  theme_bw()

# year by ovaryHolesSum
year_ovarySum <- ggeffect(logPlot, c("ovaryHolesSum [all]", "year")) %>% plot(add.data = TRUE) +
  labs(x = "Sum of Ovary Holes", y = "Predicted Prob. of Abscission",
       colour = "Year", title = 'Sum of Ovary Holes') +
  theme_bw()

# year by ovaryHolesCount
year_ovaryCount <- ggeffect(logPlot, c("ovaryHolesCount [all]", "year")) %>% plot() +
  labs(x = "Num. Ovary Holes", y = "Predicted Prob. of Abscission",
       colour = "Year", title = 'Num. Ovary Holes') +
  theme_bw()

year_ints <- arrangeGrob(year_sepal, year_bract, year_ovarySum, year_ovaryCount)
ggsave("../plots/year_grid.png", year_ints, width = 8, height = 6)
```


## Model Metrics

(make sure the models from `Modeling_Final.Rmd` have already been run and are loaded in environment)

### Decision Tree

```{r}
png('../Plots/classTree_plot.png', width = 16, height = 8, res = 96, units = "in")
rpart.plot(classTree$finalModel)
dev.off()
```

### Random Forest

```{r}
png('../Plots/RF_var_imp.png', width = 8, height = 6, res = 96, units = "in")
plot(varImp(rf))
dev.off()
```

### Boosted Trees

```{r}
png('../Plots/boost_var_imp.png', width = 8, height = 6, res = 96, units = "in")
plot(varImp(boostTree))
dev.off()
```